---
title: "Krebs"
subtitle: "Auswertungen nach regionalen und demografischen Merkmalen"
author: "Gerald Gartner"
date: "01-2019"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


```{r, echo=FALSE}
# CONFIG
user_name <- "qvvdata" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "2019-krebs" # adapt!
package_date <- "2019-01-01" # date of the CRAN snapshot that
# the checkpoint package uses
```

## Notes

This report was generated on `r Sys.time()`.

...

### R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/rscript.zip). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `ìnput` will be processed and results will be written to `output`. 
The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 

## Preparations

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work, 
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}
```


### Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(tidyverse)
library(magrittr) # pipes
library(readxl) # excel
library(jsonlite) # json
library(lintr) # code linting
library(xlsx) #Excel
library(googlesheets) # googlesheets (replace with googlesheets4 asap)",
file = "manifest.R")
```

### Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("checkpoint",
                           username = "RevolutionAnalytics",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
checkpoint(snapshotDate = package_date,
           project = path_to_wd,
           verbose = T,
           scanForPackages = T,
           use.knitr = F)
rm(package_date)
```


### Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

### Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
knitr::read_chunk("scripts/theme_addendum.R")
source("./scripts/theme_addendum.R")
source("./scripts/BorderMan.R")
source("./scripts/01-preprocessing.R")
cat(getwd())
options(scipen=999)

```




```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Österreich als Vergleichswert hinzufügen für die Bezirke je Typ/Jahr
bezirke_mortalität_at <- filter(bezirke_mortalität, wohnbezirk=="0") %>%
  select(-c(wohnbezirk, bezirkname))%>%
  rename(rate_at = rate)

# Renaming
bezirke_mortalität <- bezirke_mortalität %>%
  mutate(todesursache_kurz = ifelse(todesursache == "Bösartige Neubildungen", "alle Krebsarten", 
                                ifelse(todesursache == "Bösart. Neubild. des Magens", "Magen", 
                                       ifelse(todesursache == "Bösart. Neubild. des Colon, des Rektums und des Anus", "Darm", 
                                              ifelse(todesursache == "Bösart. Neubild. der Bauchspeicheldrüse", "Bauchspeicheldrüse", 
                                                            ifelse(todesursache == "Bösart. Neubild. der Luftröhre, Bronchien und Lunge", "Lunge", 
                                                                   ifelse(todesursache == "Bösart. Neubild. der Brustdrüse", "Brust", 
                                                                          ifelse(todesursache == "Bösart. Neubild. des Ovariums", "Eierstock", 
                                                                                 ifelse(todesursache == "Bösart. Neubild. der Prostata", "Prostata", 
                                                                                        ifelse(todesursache == "Leukämie", "Leukämie", 
                                                                                 
                                       "alle anderen")))))))))) 

# Bezirke für Recherche herausrechnen
bezirke_mortalität_vgl <- bezirke_mortalität %>%
  left_join(bezirke_mortalität_at, by=c("periode"="periode", "todesursache"="todesursache", "group"="group")) %>%
  mutate(diff_rate = rate-rate_at)  
write.xlsx(as.data.frame(bezirke_mortalität_vgl), file="output/ignore/bezirke_mortalität_vgl.xlsx", col.names=TRUE, row.names=FALSE, append=FALSE)

writegroupcsv <- function(df) {
  print(first(df$group))
  
  write_csv(df %>%
              select(-group) %>%
              gather(wertart, rate, rate, rate_at, diff_rate) %>%
              mutate(todesursache_kurz = case_when(
                todesursache_kurz=="alle Krebsarten" ~ todesursache_kurz,
                todesursache_kurz=="Leukämie" ~ todesursache_kurz,
                todesursache_kurz=="Bauchspeicheldrüse" ~ "Bauchspeicheldrüsenkrebs",
                todesursache_kurz=="Lunge" ~ "Lungenkrebs",
                TRUE ~ paste0(todesursache_kurz, "krebs")
              )) %>%
              mutate(key=paste(todesursache_kurz, wertart, sep=":")) %>%
              select(-todesursache_kurz, -wertart, -todesursache, -periode) %>%
              rename(gkz=wohnbezirk) %>%
              spread(key, rate),
            path=paste0("interaktiv/karten/dist/",first(df$group),".csv"))
  df
}

bezirke_mortalität_vgl_akt <- bezirke_mortalität_vgl %>%
  filter(periode=="2007-2015") %>%
  mutate(group=gsub("_.*$","",group))

bezirke_mortalität_vgl_akt %>%
  group_by(group) %>%
  do(writegroupcsv(.))

# Herausschreiben aller Daten für die interaktive Map
write_csv(bezirke_mortalität_vgl_akt, path="interaktiv/karten/bezirke_mortalität_vgl_akt.csv")


bezirke_mortalität_vgl_akt_map <- bezirke_mortalität_vgl_akt %>%
  filter(wohnbezirk != 0 & group =="gesamt_altersstandardisierte_rate" & todesursache_kurz=="Lunge")

bezirke_mortalität_vgl_akt_map <- bezirke_mortalität_vgl_akt_map %>% left_join(bez_18_mw, by = c("wohnbezirk" = "GKZ")) 


# visualisierung der abweichung vom bundesweiten Schnitt
plot_6 <- ggplot() +
  # municipality polygons & outline
  geom_sf(data = bezirke_mortalität_vgl_akt_map, aes(fill=diff_rate), color="white", size=0.001) +
  geom_sf(data = bundeslaendergrenzen, color="black", fill="transparent", size=0.2) +
  coord_sf() +
  #coord_map(projection = "mercator") +
  # coord_sf did not work here
  #scale_fill_gradient(low = "#ffffff", high = "#132B43",
  #                     space = "Lab", na.value = "#af3b6e", guide = "colourbar") +
  scale_fill_gradient2(low = "#84a07c", midpoint = 0, mid = "#f0edf1", high = "#ba2b58")+
  #facet_wrap(~ todesursache)+
  labs(x = NULL, y = NULL, title = "Krebs", caption = "Quelle: Statistik Austria, BEV.") +
  #guides(fill=guide_legend(nrow=6,byrow=TRUE)) +
 theme_bw()
print(plot_6)
ggsave("output/ignore/plot_6.pdf", device="pdf")

  
bezirke_mortalität_vgl_quant <- bezirke_mortalität_vgl %>%
  # select(-c(rate, rate_at)) %>%
  # mutate(periode_todesursache_group = paste(periode, todesursache, group, sep="$")) %>%
  # select(wohnbezirk, bezirkname, periode_todesursache_group, diff_rate)%>%
  # spread(periode_todesursache_group, diff_rate)
  group_by(periode, wohnbezirk, todesursache, bezirkname, group) %>%
  na.omit() %>% 
   do(data.frame(t(quantile(.$rate))))

  
# für alle Krebstypen das letzte jahr visualisieren
bezirke_mortalität_map <- bezirke_mortalität %>%
  filter(wohnbezirk != 0 & periode =="2007-2015" & group =="gesamt_altersstandardisierte_rate" & todesursache=="Bösartige Neubildungen")
  
# merge für vis
map_data <- bezirke_mortalität_map %>% left_join(bez_18_mw, by = c("wohnbezirk" = "GKZ")) 
needs(sf)
# vis
plot_1 <- ggplot() +
  # municipality polygons & outline
  geom_sf(data = map_data, aes(fill=rate), color="white", size=0.001) +
  geom_sf(data = bundeslaendergrenzen, color="black", fill="transparent", size=0.1) +
  coord_sf() +
  #coord_map(projection = "mercator") +
  # coord_sf did not work here
  scale_fill_gradient(low = "#ffffff", high = "#132B43",
                       space = "Lab", na.value = "#af3b6e", guide = "colourbar") +
  facet_wrap(~ todesursache)+
  #is_free()+
  labs(x = NULL, y = NULL, title = "Krebs", caption = "Quelle: Statistik Austria, BEV.") +
  guides(fill=guide_legend(nrow=6,byrow=TRUE)) 
print(plot_1)

```


```{r}

# Gibt es Bezirken, in denen sich die Rate negativ entwickelt?
bezirke_mortalität_trend <- bezirke_mortalität %>%
  spread(periode, rate)
  
needs(forecast)
bezirke_mortalität_trend$trend <- NA
bezirke_mortalität_trend$pvalue <- NA
for (i in 1:nrow(bezirke_mortalität_trend)){
  
  werte <- as.ts(as.vector(t(bezirke_mortalität_trend[i,][5:8])))
  fit <- tslm(werte~trend)
  trend <- fit$coefficients[2]
  pvalue <- summary(fit)$coefficients[2,4]
  
  bezirke_mortalität_trend$trend[i] <- trend
  bezirke_mortalität_trend$pvalue[i] <- pvalue
  
  cat(trend)
  cat(pvalue)
  
  
}
bezirke_mortalität_trend$sig <- NA  
bezirke_mortalität_trend$sig[bezirke_mortalität_trend$pvalue<=0.1] <- "Signifikant 5%"
bezirke_mortalität_trend$sig[bezirke_mortalität_trend$pvalue> 0.1] <- "Insignifikant 5%"
bezirke_mortalität_trend$posneg[bezirke_mortalität_trend$trend >= 0] <- "negativ"
bezirke_mortalität_trend$posneg[bezirke_mortalität_trend$trend <= 0] <- "positiv"





```

```{r}
# In wie vielen Bezirken je Bundesland liegt die Krebsmortalitätsrate über dem Bundesschnitt je Krebsart?
# LIne Chart je Krebsart je Bundesland für Entwicklung je group

# Berechnen für Oberösterreich
bezirke_mortalität_vgl_chart <- bezirke_mortalität_vgl %>%
  filter(group=="gesamt_altersstandardisierte_rate") %>%
  mutate(blcode = substr(wohnbezirk, 0, 1)) %>%
  filter(blcode==4 | blcode ==0)

plot_3 <- ggplot(bezirke_mortalität_vgl_chart, aes(x=periode, group = rev(bezirkname))) +
                   geom_line(aes(y=rate, colour=blcode), size=0.5) +
  facet_wrap(~todesursache, scales = "free") +
  scale_colour_manual(values=c("#C1839F", "#cecece"))+
  theme_bw()
                 
print(plot_3)
add_save(plot_3 + labs(title=NULL, caption=NULL),'plot_3')
#ggsave(plot_3, "output/plot_3.pdf")
```

```{r}
# Wie viele Bezirke je Bundesland sind über/unterdurchschnittlich viele Krebstote nach Typ und Rate, um zu eruieren, ob es interessante Ansätze gibt 
mortalität_yearly_tidy <- mortalität_yearly %>%
         gather(group, rate, insgesamt_absolut:frauen_kumrate)%>%
         separate(group, into = c("group", "valuetype"), sep="_") %>%
         na.omit()%>%
         mutate(rate = as.numeric(rate))

mortalität_yearly_tidy_rate <- mortalität_yearly_tidy %>% 
  filter(valuetype=="kumrate" & typ!="Gesamt")

plot_4 <- ggplot(mortalität_yearly_tidy_rate, aes(x=jahr, group = rev(typ))) +
                   geom_line(aes(y=rate), size=0.5) +
  facet_wrap(~group, scales = "free") +
  theme_bw()
print(plot_4)
```

```{r}
# Dataframe für INteractive zimmern
# 1. Step: Overall ############ praevalenz_gesamt_pct u. praevalenz_gesamt_absolut
# 2. Prävalenz ############  krebsarten_alter_pct, krebsarten_alter_absolut
# 3. Am häfuisten wird X festgesellt . Y prozent aller Diagnosen  ############ krebsarten_alter_pct_rel
#4. Zwischen X-Y Porzent überleben die nächsten 5 jahren - ueberleben_alter
# 5. Die Chance schwankt je nach Zeitpunkt der Feststellung - ueberleben_stadium
# 6. Für X in Ihrer Lebensphase hat sich die Überlebenswahrscheinlichkeit (verbessert/verschlechtert) ueberleben, ueberleben_alter

# Generate one datafile for alle cancer types
praevalenz_gesamt_pct_df <- praevalenz_gesamt_pct %>% mutate(typ = "rel")
praevalenz_gesamt_absolut_df <- praevalenz_gesamt_absolut %>% na.omit() %>% mutate(typ="abs")

krebsarten_alter_pct_df <- krebsarten_alter_pct %>% 
  mutate(lokalisation = gsub("Malignome, insgesamt \\(C00-C97, ohne C44\\)", "C99 alle-Krebsarten", lokalisation),
         typ="rel") %>% 
  separate(lokalisation, c("cs", "lokalisation"), sep = " ") %>% 
  na.omit(cancertype) %>% 
  select(-cs)


krebsarten_alter_absolut_df <- krebsarten_alter_absolut %>%
  mutate(lokalisation = gsub("Malignome, insgesamt \\(C00-C97, ohne C44\\)", "C99 alle-Krebsarten", lokalisation), 
         typ="abs") %>%
  separate(lokalisation, c("cs", "lokalisation"), sep = " ") %>% 
  na.omit(cancertype) %>% 
  select(-cs)

krebsarten_df <- bind_rows(praevalenz_gesamt_absolut_df, praevalenz_gesamt_pct_df, krebsarten_alter_absolut_df, krebsarten_alter_pct_df) %>%
  mutate(lokalisation = gsub("Gebärmutterkörper", "Gebärmutter", lokalisation)) %>%
  filter(lokalisation!="Gebärmutterhals")

write_csv(krebsarten_df, path = "interaktiv/krebsarten_df.csv")

# Generate one datafile for alle survival rates
ueberleben_alter_df <- ueberleben_alter %>% 
   separate(kre_icd10_3, c("cs", "lokalisation"), sep = ", ") %>% 
  mutate(lokalisation = gsub("Malignome", "alle-Krebsarten", lokalisation),
         typ="rel", 
         category = kre_altgrob) %>% 
  na.omit(lokalisation) %>% 
  select(-c(cs, interval, rateabove95pct, ratebelow95pct, kre_altgrob))

ueberleben_stadium_df <- ueberleben_stadium %>%
   separate(kre_icd10_3, c("cs", "lokalisation"), sep = ", ") %>% 
  mutate(lokalisation = gsub("Malignome", "alle-Krebsarten", lokalisation),
         typ="rel", 
         category = kre_stdgrob) %>% 
  na.omit(lokalisation) %>% 
  select(-c(cs, interval, rateabove95pct, ratebelow95pct, kre_stdgrob))

ueberleben_df <- bind_rows(ueberleben_stadium_df, ueberleben_alter_df)  

# Gebärmutterhals entfernen, weil nur eine Art von Gebärmutterkrebs angezeigt wird
ueberleben_df <- ueberleben_df %>%
  mutate(lokalisation = gsub("Colorectal", "Darm", lokalisation), 
         lokalisation = gsub("Gebärmutterkörper", "Gebärmutter", lokalisation)) %>%
  filter(lokalisation!="Gebärmutterhals")

# Herausschreiben der Daten für die personalisierte Grafik zum Überleben
write_csv(ueberleben_df, path = "interaktiv/ueberleben_df.csv")


```

## Linting

The code in this RMarkdown is listed with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd")
# if you have additional scripts and want them to be linted too, add them here
```


